Delayed effects of [responses to?] climate lead to demographic divergence in a fragmented Amazonian landscape

### 2021-03-24 doc/paper.Rmd
work on discussion

1.  Understanding how landscape structure and environmental conditions act to influence population dynamics is fundamental to many of the emerging conceptual frameworks for studying and conserving fragmented landscapes [@didhamRethinkingConceptualFoundations2012; @driscollConceptualDomainMatrix2013].

2.  Further evidence delayed effects may be the rule (emphasis on demographic modeling)

@@

3.  What makes our results important is that we went the next step: we not only found that there were delayed effects, but that these effects different by habitat ~~and life-history stage~~ <!--# we didn't really do this.  This implies a size * SPEI interaction -->.
    [this lays out what these patterns are]

    -   Survival at mean SPEI was similar in fragments and continuous forest for all plant sizes.
        Plants were smaller in forest fragments, but grew slightly faster than plants in continuous forest, closing the size gap over the decade of data.

    -   However, the effect of varying SPEI on survival and growth was much greater in fragments.

    -   At all sizes, variation in SPEI had a greater impact on both growth and survival.
        Both unusually wet and unusually dry conditions resulted in reductions in survival in fragments, with lagged effects of up to around 15 months.
        Similarly for plant size, extremes in both directions in the census month caused reductions in growth in fragments.
        However, the lagged effects of SPEI on plant size depended on season with the general pattern of unusually dry wet seasons and unusually wet dry seasons being beneficial to plant growth.
        Together, this could possibly be explained by reduced microclimate buffering in forest fragments.
        More research... <!--# add more, possibly move this whole paragraph. -->

@@

    -   Our results confirm those of Bruna and Kress [-@brunaHabitatFragmentationDemographic2002] showing that probability of flowering increases with plant size, and is higher in continuous forests due to the greater number of large plants compared to fragments.
        <!--# combine with first bullet point. -->

    -   Average flowering probability of plants in continuous forest is equivalent to flowering once every \_\_\_ years compared to once every \_\_\_ years in 1-ha fragments (assuming plant remains the same size and SPEI is average).
        <!--# from model intercepts -->

@@

    -   The increase in flowering rate of plants following a drought could be a bet-hedging strategy in response to stress [@nihadRelationshipStressFlowering2018].

4.  Mechanisms [this lays out the biological, fragmentation-related mechanisms underpinning these patterns]

@@

5.  Ecological Consequences of these patterns: an eye towards a drier/wetter future

>>>>

Understanding how landscape structure and environmental conditions act to influence population dynamics is fundamental to many of the emerging conceptual frameworks for studying and conserving fragmented landscapes [@didhamRethinkingConceptualFoundations2012; @driscollConceptualDomainMatrix2013].

@@

In addition to finding delayed effects of SPEI on demography, we found that these effects differed between forest fragments and continuous forest habitat.
Overall survival of plants was similar in fragments and continuous forest for all plant sizes.
Although plants were smaller in forest fragments, they grew slightly faster than plants in continuous forest, closing the size gap over the decade of demographic data.
However, at all sizes, variation in SPEI had a greater impact on both growth and survival in fragments compared to continuous forest.
Both unusually wet and unusually dry conditions resulted in reductions in survival in fragments, with lagged effects of up to around 15 months.
Similarly for plant size, extremes in both directions in the census month caused reductions in growth in fragments.
However, the lagged effects of SPEI on plant size depended on season with the general pattern of extremes (unusually wet wet seasons and unusually dry dry seasons) being detrimental to plant growth.

Our results confirm those of Bruna and Kress [-@brunaHabitatFragmentationDemographic2002] showing that probability of flowering increases with plant size, and is higher in continuous forests due to the greater number of large plants compared to fragments.
Average flowering probability of plants in continuous forest is equivalent to flowering once every 125 years compared to once every 288 years in 1-ha fragments (assuming average plant size and SPEI).
<!--# this doesn't make sense.  Would be better to pick a large plant size and plug in to both --> We also see delayed effects of drought in flowering with increased flowering in years when plants experienced a drought two dry seasons ago (15--20 months prior to census).

@@

@@

1.  Mechanisms [this lays out the biological, fragmentation-related mechanisms underpinning these patterns]

@@

2.  Ecological Consequences of these patterns: an eye towards a drier/wetter future
